---
title: "Allison, Shakira, and Gracie Final Stat 345 Project"
output: word_document
date: "2025-04-11"
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Overview:
  - We decided to look at the Top Country Songs of each decade since 1960
  - We then looked for repetition in titles and phrases over each decade
    - Looking at graphs of our findings we are able to draw conclusions about repetition

```{r student3, include= FALSE}
library(tidyverse)
#Data for all the songs
songs_data <- tibble(
  Title = c(
    "He'll Have to Go", "El Paso", "Sink the Bismarck", "Please Help Me, I'm Falling", "Let's Think About Living",
    "Something's Burning", "Snowbird", "The Wonder of You", "It's Only Make Believe", "Coward of the County",
    "Don't Fall in Love with a Dreamer", "Stand by Me", "Drivin' My Life Away", "You Decorated My Life",
    "Blaze of Glory", "Just Between You and Me", "Free Fallin'", "Breathe", "Amazed", "The Way You Love Me", "I Need You",
    "That's the Way", "Yes!", "The Best Day", "How Do You Like Me Now?!", "My Best Friend", "I Hope You Dance",
    "What About Now", "Cowboy Take Me Away", "Need You Now", "Mine", "You Belong with Me", "Stuck Like Glue", "Undo It",
    "The House That Built Me", "If I Die Young", "American Honey", "Lover, Lover", "The Bones", "I Hope", "Ten Thousand Hours",
    "Chasin' You", "One Man Band", "Got What I Got", "Die from a Broken Heart", "One of Them Girls",
    "Hard to Forget", "One Margarita", "I Hope You're Happy Now", "Even Though I'm Leavin", "Be Like That", "Homesick",
    "Bluebird", "More Than My Hometown", "Lovin' on You"
  ),
  Artist = c(
    "Jim Reeves", "Marty Robbins", "Johnny Horton", "Hank Locklin", "Bob Luman", "Kenny Rogers & The First Edition",
    "Anne Murray", "Elvis Presley", "Glen Campbell", "Kenny Rogers", "Kenny Rogers & Kim Carnes", "Mickey Gilley",
    "Eddie Rabbitt", "Kenny Rogers", "Jon Bon Jovi", "Lou Gramm", "Tom Petty", "Faith Hill",
    "Lonestar", "Faith Hill", "LeAnn Rimes", "Jo Dee Messina", "Chad Brock", "George Strait", "Toby Keith", "Tim McGraw",
    "Lee Ann Womack with Sons of the Desert", "Lonestar", "Dixie Chicks", "Lady Antebellum", "Taylor Swift", "Taylor Swift",
    "Sugarland", "Carrie Underwood", "Miranda Lambert", "The Band Perry", "Lady Antebellum", "Jerrod Niemann", "Maren Morris",
    "Gabby Barrett featuring Charlie Puth", "Dan + Shay and Justin Bieber", "Morgan Wallen", "Old Dominion", "Jason Aldean", "Maddie & Tae", "Lee Brice", "Sam Hunt", "Luke Bryan", "Carly Pearce and Lee Brice",
    "Luke Combs", "Kane Brown, Swae Lee and Khalid", "Kane Brown", "Miranda Lambert",
    "Morgan Wallen", "Luke Combs"
  ),
  Year = c(
    1959, 1959, 1960, 1960, 1960, 1970, 1970, 1970, 1970, 1979, 1980, 1980, 1980, 1979, 1990, 1989, 1989, 1999, 1999,
    2000, 2000, 2000, 2000, 2000, 1999, 1999, 2000, 2000, 1999, 2009, 2010, 2008, 2010, 2010, 2010, 2010, 2010, 2010, 2019,
    2020, 2019, 2019, 2020, 2020, 2019, 2020, 2020, 2020, 2019, 2019, 2020, 2019, 2019, 2020, 2020
  ),
  Song_Length_Seconds = c(
    140, 278, 194, 141, 122, 245, 128, 159, 145, 260, 219, 216, 197, 218, 335, 296, 254, 249, 240, 187, 228, 200, 203,
    200, 210, 279, 294, 212, 287, 277, 230, 232, 247, 177, 236, 223, 224, 198, 197, 210, 167, 205, 188, 198, 223, 178,
    193, 199, 225, 190, 205, 209, 223, 195, 205
  ),
  Word_Count = c(
    152, 466, 430, 114, 281, 191, 220, 116, 165, 419, 213, 154, 252, 161, 368, 271, 176, 254, 256, 306, 198, 467, 262,
    285, 289, 312, 318, 298, 247, 260, 406, 396, 558, 381, 297, 359, 216, 361, 326, 470, 304, 323, 315, 261, 294, 365,
    409, 311, 224, 340, 450, 276, 308, 424, 223
  ),
url = c(
  "https://genius.com/Jim-reeves-hell-have-to-go-lyrics",
  "https://genius.com/Marty-robbins-el-paso-lyrics",
  "https://genius.com/Johnny-horton-sink-the-bismarck-lyrics",
  "https://genius.com/Hank-locklin-please-help-me-im-falling-lyrics",
  "https://genius.com/Bob-luman-lets-think-about-living-lyrics",
  "https://genius.com/Kenny-rogers-and-the-first-edition-somethings-burning-lyrics",
  "https://genius.com/Anne-murray-snowbird-lyrics",
  "https://genius.com/Elvis-presley-the-wonder-of-you-lyrics",
  "https://genius.com/Glen-campbell-its-only-make-believe-lyrics",
  "https://genius.com/Kenny-rogers-coward-of-the-county-lyrics",
  "https://genius.com/Kenny-rogers-dont-fall-in-love-with-a-dreamer-lyrics",
  "https://genius.com/Mickey-gilley-stand-by-me-lyrics",
  "https://genius.com/Eddie-rabbitt-drivin-my-life-away-lyrics",
  "https://genius.com/Kenny-rogers-you-decorated-my-life-lyrics",
  "https://genius.com/Bon-jovi-blaze-of-glory-lyrics",
  "https://genius.com/Lou-gramm-just-between-you-and-me-lyrics",
  "https://genius.com/Tom-petty-free-fallin-lyrics",
  "https://genius.com/Faith-hill-breathe-lyrics",
  "https://genius.com/Lonestar-amazed-lyrics",
  "https://genius.com/Faith-hill-the-way-you-love-me-lyrics",
  "https://genius.com/Leann-rimes-i-need-you-lyrics",
  "https://genius.com/Jo-dee-messina-thats-the-way-lyrics",
  "https://genius.com/Chad-brock-yes-lyrics",
  "https://genius.com/George-strait-the-best-day-lyrics",
  "https://genius.com/Toby-keith-how-do-you-like-me-now-lyrics",
  "https://genius.com/Tim-mcgraw-my-best-friend-lyrics",
  "https://genius.com/Lee-ann-womack-i-hope-you-dance-lyrics",
  "https://genius.com/Lonestar-what-about-now-lyrics",
  "https://genius.com/Dixie-chicks-cowboy-take-me-away-lyrics",
  "https://genius.com/Lady-antebellum-need-you-now-lyrics",
  "https://genius.com/Taylor-swift-mine-lyrics",
  "https://genius.com/Taylor-swift-you-belong-with-me-lyrics",
  "https://genius.com/Sugarland-stuck-like-glue-lyrics",
  "https://genius.com/Carrie-underwood-undo-it-lyrics",
  "https://genius.com/Miranda-lambert-the-house-that-built-me-lyrics",
  "https://genius.com/The-band-perry-if-i-die-young-lyrics",
  "https://genius.com/Lady-antebellum-american-honey-lyrics",
  "https://genius.com/Jerrod-niemann-lover-lover-lyrics",
  "https://genius.com/Maren-morris-the-bones-lyrics",
  "https://genius.com/Gabby-barrett-i-hope-lyrics",
  "https://genius.com/Dan-shay-and-justin-bieber-10000-hours-lyrics",
  "https://genius.com/Morgan-wallen-chasin-you-lyrics",
  "https://genius.com/Old-dominion-one-man-band-lyrics",
  "https://genius.com/Jason-aldean-got-what-i-got-lyrics",
  "https://genius.com/Maddie-and-tae-die-from-a-broken-heart-lyrics",
  "https://genius.com/Lee-brice-one-of-them-girls-lyrics",
  "https://genius.com/Sam-hunt-hard-to-forget-lyrics",
  "https://genius.com/Luke-bryan-one-margarita-lyrics",
  "https://genius.com/Carly-pearce-and-lee-brice-i-hope-youre-happy-now-lyrics",
  "https://genius.com/Luke-combs-even-though-im-leaving-lyrics",
  "https://genius.com/Kane-brown-swae-lee-and-khalid-be-like-that-lyrics",
  "https://genius.com/Kane-brown-homesick-lyrics",
  "https://genius.com/Miranda-lambert-bluebird-lyrics",
  "https://genius.com/Morgan-wallen-more-than-my-hometown-lyrics",
  "https://genius.com/Luke-combs-lovin-on-you-lyrics"
))
```


```{r, include= FALSE}
library(dplyr)
library(rvest)
library(stringr)

# Define the scraping function 
get_lyrics_genius <- function(url) {
  tryCatch({
    h <- read_html(url)
    lyrics <- html_nodes(h, 'div[data-lyrics-container="true"]') %>%
      html_text2()
    full_lyrics <- paste(lyrics, collapse = " ")
    list(
      words = str_split(full_lyrics, "\\s+")[[1]],
      text = full_lyrics,
      lines = str_split(full_lyrics, "\\n+")[[1]]
    )
  }, error = function(e) {
    warning("Failed to scrape: ", url)
    list(words = NA, text = NA, lines = NA)
  })
}

# Add row ID to preserve alignment
songs_data <- songs_data %>% mutate(row_id = row_number())

# Pull lyrics safely
lyrics_df <- songs_data %>%
  rowwise() %>%
  mutate(
    lyrics_data = list(get_lyrics_genius(url)),
    words = list(lyrics_data$words),
    full_text = lyrics_data$text,
    lines = list(lyrics_data$lines),
    word_count = if (!is.na(lyrics_data$words[1])) length(lyrics_data$words) else NA
  ) %>%
  ungroup() %>%
  select(row_id, words, full_text, word_count, lines)

# Join lyrics back to main data
songs_df_with_lyrics <- songs_data %>%
  left_join(lyrics_df, by = "row_id") %>%
  select(-row_id)
 
```

```{r, include= FALSE}
#cleaning function applied to dataset
clean_lyrics_text <- function(full_text) {
   if (!is.na(full_text) && nchar(full_text) > 0) {
     lines <- str_split(full_text, "\n")[[1]]
     
    #CODE FOR REMOVING BEFORE BRACKET
    first_bracket_line <- which(str_detect(lines, "^\\["))[1]
    if (!is.na(first_bracket_line)) {
      lines <- lines[first_bracket_line:length(lines)]
    }
     
     # Remove any text within brackets (square, round, or curly)
     lines <- str_replace_all(lines, "\\[.*?\\]", "")  # Removes text in square brackets
     lines <- str_replace_all(lines, "\\(.*?\\)", "")  # Removes text in round brackets
     lines <- str_replace_all(lines, "\\{.*?\\}", "")  # Removes text in curly brackets
     

     # Combine the remaining lines into a clean text
     clean_text <- str_c(lines, collapse = "\n")
     
     
    # If the text starts with a number, remove everything before and the word "Lyrics"
    if (str_detect(str_trim(clean_text), "^\\d")) {
      # Look for the word "Lyrics" (case-insensitive)
      match_pos <- str_locate(tolower(clean_text), "lyrics")[1, "end"]
      if (!is.na(match_pos)) {
        clean_text <- str_sub(clean_text, match_pos + 1)
        clean_text <- str_trim(clean_text)
      }
    }
     
     return(clean_text)
   } else {
     return("")  # Return an empty string if the input is NA or empty
   }
 }

```

```{r, include= FALSE}
 #cleaning function applied to dataset
 songs_df_with_lyrics <- songs_df_with_lyrics %>%
   mutate(
     clean_full_text = map_chr(full_text, clean_lyrics_text),  
     clean_words = str_split(clean_full_text, "\\s+"),
     clean_word_count = map_int(clean_words, length)
   )
```

### Repetition in Country Songs

```{r student4, include = FALSE, warning=FALSE}
library(stringr)
library(dplyr)

# Add clean versions of both text and title to ignore random punctuation and capitalizations
songs_df_with_lyrics <- songs_df_with_lyrics %>%
  mutate(
    text_for_repetition = clean_full_text %>%
      str_to_lower() %>%
      str_replace_all("[[:punct:]]", "") %>%
      str_squish(),
    clean_title = Title %>%
      str_to_lower() %>%
      str_replace_all("[[:punct:]]", "") %>%
      str_squish()
  ) %>%
  rowwise() %>%
  mutate(
    title_count = str_count(
      text_for_repetition,
      regex(paste0("\\b", clean_title, "\\b"), ignore_case = FALSE)
    )
  ) %>%
  ungroup()

```



```{r, echo= FALSE, warning=FALSE}
#graph for average amount of times title is repeated in song by year
library(dplyr)
library(ggplot2)

# Summarize average title count per year
avg_title_per_year <- songs_df_with_lyrics %>%
  group_by(Year) %>%
  summarise(avg_title_count = mean(title_count, na.rm = TRUE))

# Plot the trend
ggplot(avg_title_per_year, aes(x = Year, y = avg_title_count)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "red", size = 2) +
  labs(
    title = "Average Title Count per Year",
    x = "Year",
    y = "Average Title Count"
  ) +
  theme_minimal()

```


Looking at the graph above, we can see a general increase in the average title usage in country songs across the decades since 1960. On the x-axis we see the years from 1960 to 2020 and on the y-axis we see the average title count in the lyrics for the top country songs. We can see notable spikes in 2000, 2010, and in 2020, which took place after large dips. We can also see 20 years of mostly stable trends from 1960 to 1980, with a peak at 1980.  Overall, there is a little trend in the average title counts over the decades with 2020's counts being much greater than 1960's. 

```{r, include= FALSE, warning=FALSE}
#function for phrase repetition 
library(stringr)
library(dplyr)
library(tidyr)

# Function to find most repeated phrase in a song
most_repeated_phrase <- function(lyrics, n = 3) {
  if (is.na(lyrics) || nchar(lyrics) == 0) return(NA)
  
  words <- str_to_lower(lyrics) %>%
    str_replace_all("[^a-z\\s]", "") %>%
    str_squish() %>%
    str_split(" ") %>%
    unlist()
  
  if (length(words) < n) return(NA)
  
  # Create n-grams
  ngrams <- sapply(1:(length(words) - n + 1), function(i) {
    paste(words[i:(i + n - 1)], collapse = " ")
  })

  # Count frequency of each phrase
  phrase_counts <- table(ngrams)
  phrase_df <- as.data.frame(phrase_counts, stringsAsFactors = FALSE)
  colnames(phrase_df) <- c("phrase", "count")
  
  # Get the most repeated phrase
  top_phrase <- phrase_df %>%
    arrange(desc(count)) %>%
    slice(1)
  
  return(top_phrase)
}

# Apply to a dataframe column
songs_df_with_lyrics <- songs_df_with_lyrics %>%
  rowwise() %>%
  mutate(top_phrase = list(most_repeated_phrase(clean_full_text)),
         most_repeated_phrase = top_phrase$phrase,
         phrase_count = top_phrase$count) %>%
  select(-top_phrase)



repetition_summary <- songs_df_with_lyrics %>%
  rowwise() %>%
  mutate(
    top_info = list(most_repeated_phrase(clean_full_text, n = 3)),  #choosing n=3 for 3 word phrases
    most_repeated_phrase = top_info$phrase,
    phrase_count = top_info$count
  ) %>%
  ungroup() %>%
  select(Title, most_repeated_phrase, phrase_count)

# Print the result
print(repetition_summary)

```

```{r, echo= FALSE, warning=FALSE}
#graph of phrase repetition across decades
library(dplyr)
library(ggplot2)

#Creating Decade column
songs_with_decade <- songs_df_with_lyrics %>%
  filter(!is.na(Year), !is.na(phrase_count)) %>%
  mutate(Decade = paste0(floor(Year / 10) * 10, "s"))

#Calculation the average phrase repetition per decade
avg_phrase_repetition <- songs_with_decade %>%
  group_by(Decade) %>%
  summarise(avg_phrase_count = mean(as.numeric(phrase_count), na.rm = TRUE)) %>%
  arrange(Decade)

#Plot
ggplot(avg_phrase_repetition, aes(x = Decade, y = avg_phrase_count, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "red", size = 2) +
  labs(
    title = "Average Phrase Repetition Across Decades",
    x = "Decade",
    y = "Average Count of Most Repeated Phrase"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


Looking at this visual, we can see that there has been an overall increase in phrase repetition in country songs since the 1950s.The x-axis is the decades, while y-axis is the average count of most popular repeated phrases. The 2000's has the highest spike of the average count of most repeated phrases in songs, which was greater than 10.0. The 2010's then still have a slight dip, but the overall trends are increasing from the 1950's to the 2020's. 


# Extending Our Findings

```{r, echo= FALSE, warning=FALSE}
library(ggplot2)
#to visualize how title count is related to word count, two y axis's are used and scaled to numbers that match the data provided
ggplot(songs_df_with_lyrics, aes(x = Year)) +
  geom_line(aes(y = title_count, color = "Title Count"), size = 1) +
  geom_line(aes(y = word_count / 10, color = "Word Count"), size = 1) +
  scale_y_continuous(
    name = "Title Count",
    limits = c(0, 100),  # controls both axes scaling
    sec.axis = sec_axis(~ . * 10, name = "Word Count", breaks = seq(0, 1000, 200))
  ) +
  scale_color_manual(values = c("Title Count" = "red", "Word Count" = "blue")) +
  labs(
    title = "Title Count vs Word Count of Each Song",
    x = "Year",
    color = "Legend"
  ) +
  theme_minimal() +
  theme(
    axis.title.y.left = element_text(color = "red"),
    axis.title.y.right = element_text(color = "blue")
  )


```


The above graph shows the count for both titles and words of country songs over time. We  can see that the word and title counts show many fluctuations. Both title and word count tend to show peaks in the same areas, which can lead us to conclude that title count may be higher for certain country songs because they have more words.There also seem to be some peaks after 2000 which could indicate an increase in word count along with title usuage in the lyrics. 

### Conclusion

For our final project, we focused on the top country songs from the 1960s to 2020. For our first repetition analysis we looked at the average title count per decade. For our second repetition analysis we looked at average phrase repetition across decades. We also extended our project to include a third graph which looks at title count vs word count for each song over time. 
For the average title count per year, we can see a general increase in title count over time, with significant peaks in the 2000s. The highest peaks in this data show over ten times for the average title count. 
For the average phrase repetition graph, we can see an overall increase from the 1960s to the 2020s. From this graph, we do see significant peaks in the 1960s and 2000s, showing an average count of repeated phrases of over ten. These peaks could reflect cultural trends and phrases that caused these peaks of phrase repetition. We can see an increase in songs being more repetitive through the second graph, through the clear trend of the phrase repetition in songs since the 1970s. 
Both graphs show an increase in activity, which can support the idea that as the repetition of song titles increases, we also see the repetition of phrases. This is also supported by the third graph, as we can see spikes in both title and word count at the corresponding times. Overall, we can see repetition has increasing trends in the top country songs from the 1960s to the 2020s. 