---
title: "Allison, Shakira, and Gracie Final Stat 345 Project"
output: word_document
date: "2025-04-11"
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

### Overview:
  - We decided to look at the Top Country Songs of each decade since 1960
  - We then looked for repetition in titles and phrases over each decade
    - Looking at graphs of our findings we are able to draw conclusion about repetition

## lyrics and songs 

```{r student3, include= FALSE}
library(tidyverse)
library(conflicted)

#Data for all the songs
songs_data <- tibble(
  Title = c(
    "He'll Have to Go", "El Paso", "Sink the Bismarck", "Please Help Me, I'm Falling", "Let's Think About Living",
    "Something's Burning", "Snowbird", "The Wonder of You", "It's Only Make Believe", "Coward of the County",
    "Don't Fall in Love with a Dreamer", "Stand by Me", "Drivin' My Life Away", "You Decorated My Life",
    "Blaze of Glory", "Just Between You and Me", "Free Fallin'", "Breathe", "Amazed", "The Way You Love Me", "I Need You",
    "That's the Way", "Yes!", "The Best Day", "How Do You Like Me Now?!", "My Best Friend", "I Hope You Dance",
    "What About Now", "Cowboy Take Me Away", "Need You Now", "Mine", "You Belong with Me", "Stuck Like Glue", "Undo It",
    "The House That Built Me", "If I Die Young", "American Honey", "Lover, Lover", "The Bones", "I Hope", "10,000 Hours",
    "Chasin' You", "One Man Band", "Got What I Got", "Die from a Broken Heart", "One of Them Girls",
    "Hard to Forget", "One Margarita", "I Hope You're Happy Now", "Even Though I'm Leaving", "Be Like That", "Homesick",
    "Bluebird", "More Than My Hometown", "Lovin' on You"
  ),
  Artist = c(
    "Jim Reeves", "Marty Robbins", "Johnny Horton", "Hank Locklin", "Bob Luman", "Kenny Rogers & The First Edition",
    "Anne Murray", "Elvis Presley", "Glen Campbell", "Kenny Rogers", "Kenny Rogers & Kim Carnes", "Mickey Gilley",
    "Eddie Rabbitt", "Kenny Rogers", "Jon Bon Jovi", "Lou Gramm", "Tom Petty", "Faith Hill",
    "Lonestar", "Faith Hill", "LeAnn Rimes", "Jo Dee Messina", "Chad Brock", "George Strait", "Toby Keith", "Tim McGraw",
    "Lee Ann Womack with Sons of the Desert", "Lonestar", "Dixie Chicks", "Lady Antebellum", "Taylor Swift", "Taylor Swift",
    "Sugarland", "Carrie Underwood", "Miranda Lambert", "The Band Perry", "Lady Antebellum", "Jerrod Niemann", "Maren Morris",
    "Gabby Barrett featuring Charlie Puth", "Dan + Shay and Justin Bieber", "Morgan Wallen", "Old Dominion", "Jason Aldean", "Maddie & Tae", "Lee Brice", "Sam Hunt", "Luke Bryan", "Carly Pearce and Lee Brice",
    "Luke Combs", "Kane Brown, Swae Lee and Khalid", "Kane Brown", "Miranda Lambert",
    "Morgan Wallen", "Luke Combs"
  ),
  Year = c(
    1959, 1959, 1960, 1960, 1960, 1970, 1970, 1970, 1970, 1979, 1980, 1980, 1980, 1979, 1990, 1989, 1989, 1999, 1999,
    2000, 2000, 2000, 2000, 2000, 1999, 1999, 2000, 2000, 1999, 2009, 2010, 2008, 2010, 2010, 2010, 2010, 2010, 2010, 2019,
    2020, 2019, 2019, 2020, 2020, 2019, 2020, 2020, 2020, 2019, 2019, 2020, 2019, 2019, 2020, 2020
  ),
  Song_Length_Seconds = c(
    140, 278, 194, 141, 122, 245, 128, 159, 145, 260, 219, 216, 197, 218, 335, 296, 254, 249, 240, 187, 228, 200, 203,
    200, 210, 279, 294, 212, 287, 277, 230, 232, 247, 177, 236, 223, 224, 198, 197, 210, 167, 205, 188, 198, 223, 178,
    193, 199, 225, 190, 205, 209, 223, 195, 205
  ),
  Word_Count = c(
    152, 466, 430, 114, 281, 191, 220, 116, 165, 419, 213, 154, 252, 161, 368, 271, 176, 254, 256, 306, 198, 467, 262,
    285, 289, 312, 318, 298, 247, 260, 406, 396, 558, 381, 297, 359, 216, 361, 326, 470, 304, 323, 315, 261, 294, 365,
    409, 311, 224, 340, 450, 276, 308, 424, 223
  ),
url = c(
  "https://genius.com/Jim-reeves-hell-have-to-go-lyrics",
  "https://genius.com/Marty-robbins-el-paso-lyrics",
  "https://genius.com/Johnny-horton-sink-the-bismarck-lyrics",
  "https://genius.com/Hank-locklin-please-help-me-im-falling-lyrics",
  "https://genius.com/Bob-luman-lets-think-about-living-lyrics",
  "https://genius.com/Kenny-rogers-and-the-first-edition-somethings-burning-lyrics",
  "https://genius.com/Anne-murray-snowbird-lyrics",
  "https://genius.com/Elvis-presley-the-wonder-of-you-lyrics",
  "https://genius.com/Glen-campbell-its-only-make-believe-lyrics",
  "https://genius.com/Kenny-rogers-coward-of-the-county-lyrics",
  "https://genius.com/Kenny-rogers-dont-fall-in-love-with-a-dreamer-lyrics",
  "https://genius.com/Mickey-gilley-stand-by-me-lyrics",
  "https://genius.com/Eddie-rabbitt-drivin-my-life-away-lyrics",
  "https://genius.com/Kenny-rogers-you-decorated-my-life-lyrics",
  "https://genius.com/Bon-jovi-blaze-of-glory-lyrics",
  "https://genius.com/Lou-gramm-just-between-you-and-me-lyrics",
  "https://genius.com/Tom-petty-free-fallin-lyrics",
  "https://genius.com/Faith-hill-breathe-lyrics",
  "https://genius.com/Lonestar-amazed-lyrics",
  "https://genius.com/Faith-hill-the-way-you-love-me-lyrics",
  "https://genius.com/Leann-rimes-i-need-you-lyrics",
  "https://genius.com/Jo-dee-messina-thats-the-way-lyrics",
  "https://genius.com/Chad-brock-yes-lyrics",
  "https://genius.com/George-strait-the-best-day-lyrics",
  "https://genius.com/Toby-keith-how-do-you-like-me-now-lyrics",
  "https://genius.com/Tim-mcgraw-my-best-friend-lyrics",
  "https://genius.com/Lee-ann-womack-i-hope-you-dance-lyrics",
  "https://genius.com/Lonestar-what-about-now-lyrics",
  "https://genius.com/Dixie-chicks-cowboy-take-me-away-lyrics",
  "https://genius.com/Lady-antebellum-need-you-now-lyrics",
  "https://genius.com/Taylor-swift-mine-lyrics",
  "https://genius.com/Taylor-swift-you-belong-with-me-lyrics",
  "https://genius.com/Sugarland-stuck-like-glue-lyrics",
  "https://genius.com/Carrie-underwood-undo-it-lyrics",
  "https://genius.com/Miranda-lambert-the-house-that-built-me-lyrics",
  "https://genius.com/The-band-perry-if-i-die-young-lyrics",
  "https://genius.com/Lady-antebellum-american-honey-lyrics",
  "https://genius.com/Jerrod-niemann-lover-lover-lyrics",
  "https://genius.com/Maren-morris-the-bones-lyrics",
  "https://genius.com/Gabby-barrett-i-hope-lyrics",
  "https://genius.com/Dan-shay-and-justin-bieber-10000-hours-lyrics",
  "https://genius.com/Morgan-wallen-chasin-you-lyrics",
  "https://genius.com/Old-dominion-one-man-band-lyrics",
  "https://genius.com/Jason-aldean-got-what-i-got-lyrics",
  "https://genius.com/Maddie-and-tae-die-from-a-broken-heart-lyrics",
  "https://genius.com/Lee-brice-one-of-them-girls-lyrics",
  "https://genius.com/Sam-hunt-hard-to-forget-lyrics",
  "https://genius.com/Luke-bryan-one-margarita-lyrics",
  "https://genius.com/Carly-pearce-and-lee-brice-i-hope-youre-happy-now-lyrics",
  "https://genius.com/Luke-combs-even-though-im-leaving-lyrics",
  "https://genius.com/Kane-brown-swae-lee-and-khalid-be-like-that-lyrics",
  "https://genius.com/Kane-brown-homesick-lyrics",
  "https://genius.com/Miranda-lambert-bluebird-lyrics",
  "https://genius.com/Morgan-wallen-more-than-my-hometown-lyrics",
  "https://genius.com/Luke-combs-lovin-on-you-lyrics"
))
```


```{r, include= FALSE}
library(dplyr)
library(rvest)
library(stringr)
library(purrr)
  
songs_data <- songs_data %>% mutate(row_id = row_number())

 # Pull lyrics safely
 lyrics_df <- songs_data %>%
   rowwise() %>%
   mutate(
     lyrics_data = list(get_lyrics_genius(url)),
     words = list(lyrics_data$words),         # wrap in list()
     full_text = lyrics_data$text,
     word_count = if (!is.na(lyrics_data$words[1])) length(lyrics_data$words) else NA
   ) %>%
   ungroup() %>%
   select(row_id, words, full_text, word_count)
 
 # Join lyrics back to main data
 songs_df_with_lyrics <- songs_data %>%
   left_join(lyrics_df, by = "row_id") %>%
   select(-row_id)
 
```

```{r, include= FALSE}
#cleaning function applied to dataset
clean_lyrics_text <- function(full_text) {
   if (!is.na(full_text) && nchar(full_text) > 0) {
     lines <- str_split(full_text, "\n")[[1]]
     
    #CODE FOR REMOVING BEFORE BRACKET
    first_bracket_line <- which(str_detect(lines, "^\\["))[1]
    if (!is.na(first_bracket_line)) {
      lines <- lines[first_bracket_line:length(lines)]
    }
     
     # Remove any text within brackets (square, round, or curly)
     lines <- str_replace_all(lines, "\\[.*?\\]", "")  # Removes text in square brackets
     lines <- str_replace_all(lines, "\\(.*?\\)", "")  # Removes text in round brackets
     lines <- str_replace_all(lines, "\\{.*?\\}", "")  # Removes text in curly brackets
     

     # Combine the remaining lines into a clean text
     clean_text <- str_c(lines, collapse = "\n")
     
     
    # If the text starts with a number, remove everything before and the word "Lyrics"
    if (str_detect(str_trim(clean_text), "^\\d")) {
      # Look for the word "Lyrics" (case-insensitive)
      match_pos <- str_locate(tolower(clean_text), "lyrics")[1, "end"]
      if (!is.na(match_pos)) {
        clean_text <- str_sub(clean_text, match_pos + 1)
        clean_text <- str_trim(clean_text)
      }
    }
     
     return(clean_text)
   } else {
     return("")  # Return an empty string if the input is NA or empty
   }
 }

```

```{r, include= FALSE}
 #cleaning function applied to dataset
 songs_df_with_lyrics <- songs_df_with_lyrics %>%
   mutate(
     clean_full_text = map_chr(full_text, clean_lyrics_text),  
     clean_words = str_split(clean_full_text, "\\s+"),
     clean_word_count = map_int(clean_words, length)
   )
```



## Section Four (Working together on function)

```{r student4, include = FALSE}
#function for title repititon
songs_df_with_lyrics <- songs_df_with_lyrics %>%
  rowwise() %>%
  mutate(
    title_count = str_count(
    str_to_lower(clean_full_text), 
    regex(paste0("\\b", str_to_lower(Title), "\\b"))
    )
  ) %>%
  ungroup()

songs_df_with_lyrics %>% select(Title, title_count)

```

```{r, echo= FALSE}
#graph for average amount of times title is repeated in song by year
library(dplyr)
library(ggplot2)

# Summarize average title count per year
avg_title_per_year <- songs_df_with_lyrics %>%
  group_by(Year) %>%
  summarise(avg_title_count = mean(title_count, na.rm = TRUE))

# Plot the trend
ggplot(avg_title_per_year, aes(x = Year, y = avg_title_count)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "darkred", size = 2) +
  labs(
    title = "Average Title Count per Year",
    x = "Year",
    y = "Average Title Count"
  ) +
  theme_minimal()

```
Looking at the graph above, we can see little trend in the average title usuage in country songs across the decads sinces 1960.

```{r, include= FALSE}
#function for phrase repetition (I pasted this from chat so it needs to be looked at)
library(stringr)
library(dplyr)
library(tidyr)

# Function to find most repeated phrase in a song
most_repeated_phrase <- function(lyrics, n = 3) {
  if (is.na(lyrics) || nchar(lyrics) == 0) return(NA)
  
  # Clean and tokenize lyrics
  words <- str_to_lower(lyrics) %>%
    str_replace_all("[^a-z\\s]", "") %>%
    str_squish() %>%
    str_split(" ") %>%
    unlist()
  
  if (length(words) < n) return(NA)
  
  # Create n-grams
  ngrams <- sapply(1:(length(words) - n + 1), function(i) {
    paste(words[i:(i + n - 1)], collapse = " ")
  })

  # Count frequency of each phrase
  phrase_counts <- table(ngrams)
  phrase_df <- as.data.frame(phrase_counts, stringsAsFactors = FALSE)
  colnames(phrase_df) <- c("phrase", "count")
  
  # Get the most repeated phrase
  top_phrase <- phrase_df %>%
    arrange(desc(count)) %>%
    slice(1)
  
  return(top_phrase)
}

# Apply to a dataframe column
songs_df_with_lyrics <- songs_df_with_lyrics %>%
  rowwise() %>%
  mutate(top_phrase = list(most_repeated_phrase(clean_full_text)),
         most_repeated_phrase = top_phrase$phrase,
         phrase_count = top_phrase$count) %>%
  select(-top_phrase)



repetition_summary <- songs_df_with_lyrics %>%
  rowwise() %>%
  mutate(
    top_info = list(most_repeated_phrase(clean_full_text, n = 3)),  # Change n for bigram/trigram
    most_repeated_phrase = top_info$phrase,
    phrase_count = top_info$count
  ) %>%
  ungroup() %>%
  select(Title, most_repeated_phrase, phrase_count)

# Print the result
print(repetition_summary)

```

```{r, echo= FALSE}
#graph of phrase repetition across decades (this is also from chat)
library(dplyr)
library(ggplot2)

#Creating Decade column
songs_with_decade <- songs_df_with_lyrics %>%
  dplyr::filter(!is.na(Year), !is.na(phrase_count)) %>%
  mutate(Decade = paste0(floor(Year / 10) * 10, "s"))

#Calculation the average phrase repetition per decade
avg_phrase_repetition <- songs_with_decade %>%
  group_by(Decade) %>%
  summarise(avg_phrase_count = mean(as.numeric(phrase_count), na.rm = TRUE)) %>%
  arrange(Decade)

#Plot
ggplot(avg_phrase_repetition, aes(x = Decade, y = avg_phrase_count, group = 1)) +
  geom_line(color = "black", size = 1) +
  geom_point(color = "red", size = 2) +
  labs(
    title = "Average Phrase Repetition Across Decades",
    x = "Decade",
    y = "Average Count of Most Repeated Phrase"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
Looking at this visual, we can see that there has been an overall increase in phrase repetition in country songs since the 1950s.The 2000's has the highest spike of the average count of most repeated phrases in songs, which was greater than 10.0. 

### Conclusion:
  - 

